---
title: "using tidymodels to model vaccination rate data"
output: github_document
---

### setup

```{r message = F, warning = F}
tictoc::tic()
library(tidyverse)
library(tidymodels)
library(rstanarm)

theme_custom = tvthemes::theme_avatar() +
  theme(plot.title = element_text(hjust = 0.5),
        panel.grid.major = element_line(linewidth = 0.5, colour = "#D6D0C4"),
        panel.grid.minor = element_line(linewidth = 0.5, colour = "#D6D0C4"))

theme_set(theme_custom)
```

___

### data import

```{r}
link = "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-25/measles.csv"
measles = read_csv(link, col_types = cols())
glimpse(measles)
```

___

### creating data frame tracking if schools are above or below threshold

```{r}
measles_df = measles |>
  filter(mmr > 0) |>
  transmute(state,
            mmr_threshold = case_when(mmr > 95 ~ "above",
                                      T ~ "below")) |>
  mutate_if(is.character, factor)

sample_n(measles_df, 10)
```

___

### average rate of vaccination by state

```{r}
measles_df |>
  group_by(state) |>
  summarise(mmr = round(mean(mmr_threshold == "above"), 3)) |>
  arrange(desc(mmr)) |>
  ggplot(aes(reorder(state, mmr), mmr)) +
  geom_col(fill = "#A390AD", position = "dodge") +
  geom_text(aes(label = mmr), hjust = -0.25, size = 3) +
  coord_flip(ylim = c(0, 0.95)) +
  labs(x = "state", y = "above threshold rate",
       title = "average vaccination rate by state")
```

___

### above/below threshold counts by state

```{r}
measles_df |>
  count(state, mmr_threshold) |>
  ggplot(aes(reorder(state, n), n)) +
  geom_col(aes(fill = mmr_threshold), position = "dodge") +
  coord_flip() +
  scale_fill_manual(values = c("#A5BDA3", "#D89E9E")) +
  labs(x = "state", y = "count", fill = "above/below threshold",
       title = "counts of schools above/below threshold by state") +
  theme(legend.position = "right")
```

___

### fitting a basic logistic regression model

```{r}
glm_fit = logistic_reg() |>
  set_engine("glm") |>
  fit(mmr_threshold ~ state, data = measles_df)

glm_fit
```

___

### getting predictions and confident intervals

```{r}
new_schools = tibble(state = unique(measles_df$state))
mean_pred = predict(glm_fit, new_data = new_schools, type = "prob")
conf_int = predict(glm_fit, new_data = new_schools, type = "conf_int")

schools_res = new_schools |>
  bind_cols(mean_pred) |>
  bind_cols(conf_int)

glimpse(schools_res)
```

___

### viewing modeling results

```{r}
schools_res |>
  mutate(state = fct_reorder(state, .pred_above)) |>
  ggplot(aes(state, .pred_above)) +
  geom_col(aes(fill = state), show.legend = F, alpha = 0.75) +
  geom_errorbar(aes(ymin = .pred_lower_above, ymax = .pred_upper_above), alpha = 0.5) +
  scale_y_continuous(labels = scales::percent_format()) +
  coord_flip() +
  labs(y = "predicted probability of being above threshold",
       title = "modeling results")
```

___

### fitting a Bayesian model

```{r}
options(mc.cores = parallel::detectCores())
prior_dist = student_t(df = 2)

# this takes a while to finish up
stan_fit = logistic_reg() |>
  set_engine("stan",
             prior = prior_dist,
             prior_intercept = prior_dist) |>
  fit(mmr_threshold ~ state, data = measles_df)

stan_fit
```

___

### getting modeling results in a data frame

```{r}
bayes_pred = predict(stan_fit,
                     new_data = new_schools,
                     type = "prob")

bayes_int = predict(stan_fit,
                    new_data = new_schools,
                    type = "conf_int")

bayes_res = new_schools |>
  bind_cols(bayes_pred) |>
  bind_cols(bayes_int)

glimpse(bayes_res)
```

___

### comparing basic logistic regression and Bayesian modeling results

```{r}
schools_res |>
  mutate(model = "glm") |>
  bind_rows(bayes_res |>
              mutate(model = "stan")) |>
  mutate(stat = fct_reorder(state, .pred_above)) |>
  ggplot(aes(state, .pred_above)) +
  geom_point(aes(col = model), size = 4) +
  geom_errorbar(aes(ymin = .pred_lower_above, ymax = .pred_upper_above, col = model),
                linewidth = 1.5, alpha = 0.75, position = "dodge") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_color_manual(values = c("#7AAC82", "#B38FB3")) +
  coord_flip() +
  labs(y = "predicted probability of being above threshold",
       title = "comparing modeling results")
```

___

### script runtime

```{r}
tictoc::toc()
```






















































